Do return series have power-law tails with the same index?

### 2017-07-21 papers/FX/draft.tex
Added my own thesis.

% \documentclass{article}
% \usepackage{amsmath}
% \usepackage{amsthm}
% \usepackage{enumerate}
% \usepackage{bookmark}
% \usepackage{graphicx}
% \usepackage{multirow}
% \usepackage{adjustbox}
% \usepackage{wrapfig}
% \usepackage{subcaption}
% \usepackage{color}
% \usepackage[dvipsnames]{xcolor}

% \usepackage{amssymb,amsmath,amsthm,amsfonts}
% \usepackage{mathrsfs}
% \usepackage{dsfont}
% \usepackage{enumerate}

% \input{physics_common.tex}

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
%% \usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

\usepackage{amssymb,amsmath,amsthm,amsfonts}
%\usepackage{mathrsfs}
%\usepackage{dsfont}
%\usepackage{enumerate}

%\input{../../physics_common.tex}
%\input{physics_common.tex}

\textwidth 6.50in
\topmargin -0.50in
\oddsidemargin 0in
\evensidemargin 0in
\textheight 9.00in
%\pagestyle{plain}
\definecolor{darkblue}{rgb}{.2, 0.2,.8}
\definecolor{darkgreen}{rgb}{0,0.5,0.3}
\definecolor{darkred}{rgb}{.8, .1,.1}
\newcommand{\pd}{\partial}
\newcommand{\red}{\color{darkred}}
\newcommand{\blue}{\color{darkblue}}
\newcommand{\green}{\color{darkgreen}}
\newcommand{\black}{\color{black}}
\newcommand{\levy}{L\'evy}
\newcommand{\slln}{strong law of large numbers}
\newcommand{\clt}{central limit theorem}
\newcommand{\sde}{stochastic \de}
\newcommand{\de}{differential equation}
\newcommand{\mregvar}{multivariate \regvar}
\newcommand{\garch}{{\rm GARCH}$(1,1)$}
\newcommand{\spr}{stochastic process}
\newcommand{\It}{It\^o}
\newcommand{\sta}{St\u aric\u a}
\newcommand{\ex}{{\rm e}\,}
\newcommand{\sas}{s$\alpha$s}
\newcommand{\civ}{\stackrel{v}{\rightarrow}}
\newcommand{\asy}{asymptotic}
\newcommand {\hatt} {\hat{\theta}}
\newcommand {\hattg} {\hat{\theta}_{MLE}}
\newcommand {\hattz} {\hat{\theta}_0}
\newcommand {\hatb} {\hat{\beta}}
\newcommand {\hatbg} {\hat{\beta}_G}
\newcommand {\spp} {^\prime}
\newcommand{\ts}{time series}
\newcommand{\tsa}{\ts\ analysis}
\newcommand{\garchpq}{{\rm GARCH}$(p,q)$}
\def\theequation{\thesection.\arabic{equation}}
\def\tag{\refstepcounter{equation}\leqno }
\def\neqno{\refstepcounter{equation}\leqno(\thesection.\arabic{equation})}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{figur}[lemma]{Figure}
\newtheorem{theorem}[lemma]{Theorem}
\newcommand{\Leb}{{\mathbb L}{\mathbb E}{\mathbb B}}
\newcommand{\bbc}{{\mathbb C}}
\newtheorem{proposition}[lemma]{Proposition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{example}[lemma]{Example}
\newtheorem{exercise}[lemma]{Exercise}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{fig}[lemma]{Figure}
\newtheorem{tab}[lemma]{Table}
\newtheorem{hypothesis}{Hypothesis}
\newcommand{\play}{\displaystyle}
\newcommand{\rmk}{\noindent R{\sc emark}}
\newcommand{\cid}{\stackrel{d}{\rightarrow}}
\newcommand{\cip}{\stackrel{P}{\rightarrow}}
\newcommand{\reference}{\hangindent=5pc\hangafter=1}
\newcommand{\MC}{Markov chain}
\newcommand{\bfQ}{{\bf Q}}
\newcommand{\bfM}{{\bf M}}
\newcommand{\bfu}{{\bf u}}
\newcommand{\bfv}{{\bf v}}
\newcommand{\bfV}{{\bf V}}
\newcommand{\bfU}{{\bf U}}
\newcommand{\bfR}{{\bf R}}
\newcommand{\bfC}{{\bf C}}
\newcommand{\bfF}{{\bf F}}
\newcommand{\RV}{{\rm RV}}
\newcommand{\LN}{{\rm LN}}
\newcommand{\WE}{{\rm WE}}
\newcommand{\bth}{\begin{theorem}}
\newcommand{\ethe}{\end{theorem}}
\newcommand{\sv}{stochastic volatility}
\newcommand{\bre}{\begin{remark}\em }
\newcommand{\ere}{\end{remark}}
\newcommand{\fre}{frequenc}
\newcommand{\ble}{\begin{lemma}}
\newcommand{\ele}{\end{lemma}}
\newcommand{\sre}{stochastic recurrence equation}
\newcommand{\pp}{point process}
\newcommand{\bde}{\begin{definition}}
\newcommand{\ede}{\end{definition}}
\newcommand{\bco}{\begin{corollary}}
\newcommand{\eco}{\end{corollary}}

\newcommand{\bpr}{\begin{proposition}}
\newcommand{\epr}{\end{proposition}}

\newcommand{\bexer}{\begin{exercise}}
\newcommand{\eexer}{\end{exercise}}

\newcommand{\bexam}{\begin{example}}
\newcommand{\eexam}{\end{example}}

\newcommand{\bfi}{\begin{fig}}
\newcommand{\efi}{\end{fig}}

\newcommand{\btab}{\begin{tab}}
\newcommand{\etab}{\end{tab}}
%\renewcommand{\thetab}{\arabic{section}.9.\arabic{tab}}

\newcommand{\edf}{empirical distribution function}
\newcommand{\epro}{empirical process}
\newcommand{\per}{periodogram}
\newcommand{\pf}{{\bf Proof.}}
\newcommand{\lhs}{left-hand side}
\newcommand{\fidi}{finite-dimensional distribution}
\newcommand{\rv}{random variable}
\newcommand{\DA}{{\rm DA}}
\newcommand{\DNA}{{\rm DNA}}
\newcommand{\MDA}{{\rm MDA}}
\newcommand{\sign}{{\rm sign}}
\newcommand{\PRM}{{\rm PRM}}
\newcommand{\card}{{\rm card}}
\newcommand{\var}{{\rm var}}
\newcommand{\med}{{\rm med}}
\newcommand{\cov}{{\rm cov}}
\newcommand{\corr}{{\rm corr}}
\newcommand{\as}{{\rm a.s.}}
\newcommand{\io}{{\rm i.o.}}
\newcommand{\Pois}{\rm Poisson}

\newcommand{\bfth}{\mbox{\boldmath$\theta$}}
\newcommand{\bfTh}{\mbox{\boldmath$\Theta$}}
\newcommand{\bfmu}{\mbox{\boldmath$\mu$}}

\newcommand{\bfvep}{\mbox{\boldmath$\vep$}}
\newcommand{\rhs}{right-hand side}
\newcommand{\df}{distribution function}
\newcommand{\dsum}{\displaystyle\sum}
%\newcommand{\dfrac}{\displaystyle\frac}
\newcommand{\dint}{\displaystyle\int}
\newcommand{\dlim}{\displaystyle\lim\limits}
\newcommand{\dmax}{\displaystyle\max\limits}
\newcommand{\dsup}{\displaystyle\sup\limits}
\newcommand{\dmin}{\displaystyle\min\limits}
\newcommand{\dprod}{\displaystyle\prod}

\newcommand{\beao}{\begin{eqnarray*}}
\newcommand{\eeao}{\end{eqnarray*}\noindent}

\newcommand{\beam}{\begin{eqnarray}}
\newcommand{\eeam}{\end{eqnarray}\noindent}

\newcommand{\beqq}{\begin{equation}}
\newcommand{\eeqq}{\end{equation}\noindent}

\newcommand{\bce}{\begin{center}}
\newcommand{\ece}{\end{center}}

\newcommand{\barr}{\begin{array}}
\newcommand{\earr}{\end{array}}
\newcommand{\cadlag}{c\`adl\`ag}
\newcommand{\tep}{\tilde\epsilon}
\newcommand{\stp}{\stackrel{\P}{\rightarrow}}
\newcommand{\std}{\stackrel{d}{\rightarrow}}
\newcommand{\stas}{\stackrel{\rm a.s.}{\rightarrow}}
\newcommand{\stj}{\stackrel{J_1}{\rightarrow}}
\newcommand{\stv}{\stackrel{v}{\rightarrow}}
\newcommand{\stw}{\stackrel{w}{\rightarrow}}
\newcommand{\stww}{\stackrel{\wh w}{\rightarrow}}
\newcommand{\eqd}{\stackrel{d}{=}}

\newcommand{\refbib}{\cite}
\newcommand{\vague}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
                    \it{v} $}}{\rightarrow}}
\newcommand{\weak}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
                    \it{w} $}}{\rightarrow}}
\newcommand{\what}{\stackrel{\lower0.2ex\hbox{$\scriptscriptstyle
                    \it{\hat{w}} $}}{\rightarrow}}

\newcommand{\bdis}{\begin{displaymath}}
\newcommand{\edis}{\end{displaymath}\noindent}

\newcommand{\Thetabold}{{\boldsymbol{\Theta}}}
\newcommand{\Prob}{\operatorname{P}}
\newcommand{\Dspace}{\mathbb{D}}
\newcommand{\Espace}{\mathbb{E}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Xbold}{{\mathbf{X}}}
\newcommand{\xbold}{{\mathbf{x}}}
\newcommand{\Ybold}{{\mathbf{Y}}}
\newcommand{\ybold}{{\mathbf{y}}}
\newcommand{\zerobold}{{\mathbf{0}}}
\newcommand{\Sphere}{\mathbb{S}}
\newcommand{\Dbar}{\overline{\mathbb{D}}_{0}}
\newcommand{\ind}{independent}
\newcommand{\w}{\omega}
\newcommand{\W}{\Omega}
\newcommand{\bw}{\bigwedge}
\newcommand{\bv}{\bigvee}
\newcommand{\nto}{n\to\infty}
\newcommand{\kto}{k\to\infty}
\newcommand{\xto}{x\to\infty}
\newcommand{\tto}{T\to\infty}
\newcommand{\uto}{u\to\infty}
\newcommand{\ov}{\overline}
\newcommand{\wt}{\widetilde}
\newcommand{\wh}{\widehat}
\newcommand{\vep}{\varepsilon}
\newcommand{\ep}{\epsilon}
\newcommand{\vap}{\varphi}
\newcommand{\la}{\lambda}
\newcommand{\La}{\Lambda}

\newcommand{\halmos}{\hfill \qed}
\newcommand{\regvary}{regularly varying}
\newcommand{\slvary}{slowly varying}
\newcommand{\regvar}{regular variation}
%\newcommand{\bbr}{{I\!\!R}}
%\newcommand{\bbn}{{I\!\!N}}
%\newcommand{\bbz}{{I\!\!\!Z}}

\newcommand{\ka}{\kappa}
\newcommand{\bba}{{\mathcal A}}
\newcommand{\bbe}{{\ov \bbr^{d}_0}}
\newcommand{\bfT}{{\bf T}}
\newcommand{\bbr}{{\mathbb R}}
\newcommand{\bbm}{{\mathcal M}}
\newcommand{\bbb}{{\mathcal B}}
\newcommand{\bbz}{{\mathbb Z}}
\newcommand{\bbn}{{\mathbb N}}
\newcommand{\bbd}{{\mathbb D}}
\newcommand{\bbf}{{\mathcal F}}
\newcommand{\bbs}{{\mathbb S}}
\newcommand{\bbE}{{\mathcal E}}

\newcommand{\BM}{Brownian motion}
\newcommand{\BB}{Brownian bridge}
\newcommand{\con}{convergence}
\newcommand{\ev}{extreme value}
\newcommand{\evt}{extreme value theory}
\newcommand{\evd}{extreme value distribution}
\newcommand{\clp}{central limit problem}
\newcommand{\nc}{norming constant}
\newcommand{\st}{such that}
\newcommand{\fif}{if and only if}
\newcommand{\wrt}{with respect to}
\newcommand{\chf}{characteristic function}
\newcommand{\fct}{function}
\newcommand{\mdoa}{maximum domain of attraction}
\newcommand{\ds}{distribution}
\newcommand{\gev}{generalized extreme value distribution}
\newcommand{\gpd}{generalized Pareto distribution}
\newcommand{\pwm}{probability-weighted moments}
\newcommand{\mle}{maximum likelihood estimator}
\newcommand{\mel}{mean excess function}
\newcommand{\an}{asymptotically normal}
\newcommand{\ac}{absolutely continuous}
\newcommand{\rep}{representation}
\newcommand{\cmt}{continuous mapping theorem}
\newcommand{\seq}{sequence}
\newcommand{\pr} {{\bf Proof.}}
\newcommand{\ins}{insurance}
\newcommand{\mat}{mathematic}
\newcommand{\pro}{probabilit}
\newcommand{\pros}{probabilities}
\newcommand{\ms}{measure}
\newcommand{\mgf}{moment generating function}
\newcommand{\ld}{large deviation}
\newcommand{\bfx}{{\bf x}}
\newcommand{\bfX}{{\bf X}}
\newcommand{\bfW}{{\bf W}}
\newcommand{\bfB}{{\bf B}}
\newcommand{\bfY}{{\bf Y}}
\newcommand{\bfy}{{\bf y}}
\newcommand{\bfA}{{\bf A}}
\newcommand{\bfG}{{\bf G}}
\newcommand{\bfO}{{\bf 0}}
\newcommand{\bfZ}{{\bf Z}}
\newcommand{\bfz}{{\bf z}}
\newcommand{\bfS}{{\bf S}}
\newcommand{\bfa}{{\bf a}}
\newcommand{\bfb}{{\bf b}}
\newcommand{\bfe}{{\bf e}}
\newcommand{\bft}{{\bf t}}
\newcommand{\bfI}{{\bf I}}
\newcommand{\bfs}{{\bf s}}
\newcommand{\bfh}{{\bf h}}
\newcommand{\bfunit}{{\bf h}}
\newcommand{\ARCH}{{\rm ARCH}}
\newcommand{\E }{{\mathbb E}}
\renewcommand{\P }{{\mathbb P}}
\newcommand{\Q }{{\mathbb Q}}
\newcommand{\1}{{\mathbf 1}}

\newcommand{\floor}[1]{
  \lfloor #1\rfloor
}
\newcommand{\ceil}[1]{
  \lceil #1\rceil
}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\allowdisplaybreaks
\begin{document}
\title{Do return series have power-law tails with the same index?}
t\author{Thomas Mikosch,  Casper de Vries, Xie Xiaolei}
\date{\today}

@@

\begin{abstract}
%We consider an investor with preferences that accord with Generalized
%Disappointment Aversion. Such an investor cares about downside risk and
%we assume he recognizes the heavy tail feature of asset return
%distributions. We argue that when a market is dominated by rational
%investors of this kind, the return distributions of equities that are
%actively traded in this market must have very similar tail-indices. We
%also show, in contrast, the scale parameters of the return
%distributions may differ hugely from one another.
%On the other hand, whether or not all equities in a multivariate model
%have the same tail-index is a dividing issue for multivariate GARCH
%models proposed in the literature. Therefore, it is important to analyze
%data of real equity returns and see how close to each other the
%tail-indices actually are.
%In this work empirical results are also presented and they appear
%to support the conclusion that the tail-indices are very similar,
%with respect to the confidence bands of estimation.
\end{abstract}

\section{Introduction}\setcounter{equation}{0}
It is one of the stylized facts of financial econometrics that 
returns of speculative prices are {\em heavy-tailed.} There is no agreement
in the literature about how heavy these tails really are. For example, Barndorff-Nielsen and Shephard
\cite{barndorff:shephard:2001} and Eberlein \cite{eberlein:2001} favor ``semi-heavy'' tails
which are comparable with those of a gamma \ds . On the other
hand, tails of returns have been studied in great detail 
in the extreme value community. Among extreme value specialists
there is general agreement that returns $X_t$ have tails of power-law-type, i.e.,
\beam\label{eq:1}
\P(X_t>x)\sim c_+ \,x^{-\alpha_{\rm up}}\quad\mbox{and}\quad \P(X_t<-x)\sim c_-\,x^{-\alpha_{\rm low}}\,,\quad\xto\,,
\eeam  
where $c_{\pm}$, $\alpha_{\rm up}$ and $\alpha_{\rm low}$ are positive constants.\footnote{Here and in what
  follows, $f(x)\sim g(x)$ for positive \fct s $f$ and $g$ means 
that $f(x)/g(x)\to 1$ as $\xto$.} See for example, Embrechts et
al. \cite{embrechts:klueppelberg:mikosch:1997}, Jansen and de Vries
\cite{jansen1991frequency}, Mikosch \cite{mikosch:2003}, Resnick
\cite{resnick:2007}.  
In the extreme value literature it is common to replace the constants $c_\pm$ by suitable
{\em slowly varying} \fct s; cf. Embrechts et
al. \cite{embrechts:klueppelberg:mikosch:1997}, Chapter 3. In this
paper, for the  
sake of argument, we stick to the condition \eqref{eq:1}.
\par
There are some good theoretical reasons for the appearance of power-law tails in situations where certain moments 
of data are believed to be infinite. Tails of type \eqref{eq:1}
describe the maximum domain of attraction of the  
Fr\'echet \ds\ $\Phi_{\alpha_{\rm up}}(x)=\exp(-x^{-\alpha_{\rm up}})$ for $x>0$, i.e., scaled maxima of an iid \seq\ $(X_t)$ with
upper tail described in \eqref{eq:1} converge in \ds\ to $\Phi_{\alpha_{\rm up}}$.
Equivalently, power-law tails are prescribed by the 
generalized Pareto \ds\ which is the limit \ds\ of the excesses of $X_t$ above high thresholds, i.e., for a suitable positive scaling \fct\ $a(u)$,
\beao
\P((X_t-u)/a(u)>x \mid X_t>x) \to (1+ x/\alpha_{\rm up})^{-\alpha_{\rm up}}\,,\qquad u\to\infty\,.
\eeao
The aforementioned results 
are considered very natural for iid and weakly dependent strictly stationary \seq s of random variables $(X_t)$; 
in the world of extremes they are the analogs
of the \clt\ from the world of sums.
\par
In the literature on extremes for return data one finds the 
statement that {\em estimated} values $\hat \alpha_{\rm up}$ and $\hat \alpha_{\rm low}$ 
of the tail-indices $\alpha_{\rm up}$ and $\alpha_{\rm low}$,
respectively, typically have the tendency that $\hat \alpha_{\rm
  up}>\hat \alpha_{\rm low}$. 
This observation is often explained by the fact that investors are
more prone to negative than to positive news in the market. 
Moreover, in the literature the {\em estimated} tail-indices $\hat \alpha$ (both in the left and right tails) 
are typically found in the range $(2,4)$. For an illustration, see
Figure~\ref{fig:1} where estimates $\hat \alpha_{\rm low}$ 
in three sectors of the Standard \& Poors 500 index  are shown. The
estimates are based on 1304 observations of daily return 
data from 4 January 2010 to 31 December 2014.
\par
When looking at Figure~\ref{fig:1} one might ask the following questions:
\begin{itemize}
\item
In view of the wide \asy\ confidence bands for the estimators of tail-indices, 
are the tail-indices from different series really distinct?
\item
Are there some {\em theoretical} reasons supporting the fact that the tail-indices from different series are {\em not} 
distinct?
\end{itemize}
In this paper, we try to find some answers to these questions. 
\par
The estimator of the tail-index $\alpha>0$  in the model
\beao
\P(X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeao
favored in the literature  is the {\em Hill estimator}; the graphs in 
Figure~\ref{fig:1} are based on this estimator. We introduce this estimator in Section~\ref{sec:1} and discuss 
some of its virtues and vices. In addition to tail-index estimation we also discuss the related problem of
estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). 
%{\red check} It turns out 
%that the simultaneous estimation of the tail-index and the scale parameter are strongly related.
\par
In Section~\ref{sec:2} we discuss the theoretical problem of appearance of power-law tails in models 
for daily or, more generally, low-\fre y 
return data. In particular, in Section~\ref{subsec:garch} we address the power-law tails of univariate and multivariate GARCH models as potential models
for a set of return data from distinct assets. As a matter of fact, under mild conditions, the aforementioned models
have power-law tails due to their relation with so-called {\em \sre s}. Moreover, some of the {\em standard} multivariate 
GARCH models as the CCC ensure that the component-wise marginal \ds s have power-law tails with the same index.
\par
In Section~\ref{sec:3} we discuss an economic argument for the fact that return data of similar assets
(such as return series in a given sector of the S\&P 500 index) may have tail-indices which are close to each other.
We argue based on a  utility \fct\ approach. We explicitly recognize the behavioral
concern for downside risk in an investor's evaluation of a portfolio
using the framework of Generalized Disappointment Aversion (GDA)
introduced by Routledge and Zin \cite{routledge2010generalized}. GDA
is an extension of the concept of Disappointment Aversion (DA) of Gul \cite{gul:1991} who derived DA from first principles (axiomatic).

In Section~\ref{sec:4} we summarize the discussion of the previous
sections.

\section{Power-law tails of return series: some empirical results}\label{sec:1}\setcounter{equation}{0}
In this section, we assume the model \eqref{eq:1} for the tails of the marginal
\ds\ of a univariate return series $(X_t)$. For the sake of argument, we assume
that this series constitutes a strictly stationary \seq . In what follows, we focus
on the left tail of the \ds , i.e., on the losses. 
However, it is common to present the tail-index estimators
for positive data. Therefore we will multiply the losses $X_t$ by
minus one, swapping the negative with the positive values.
For simplicity, we also suppress subscripts in the notation:
\beam\label{eq:1a}
\P(-X_t>x)\sim c\,x^{-\alpha}\,,\qquad \xto\,,
\eeam
where we assume that the two parameters -- the tail-index  $\alpha$ and the scale parameter
$c$ -- are positive. They play crucial roles for the understanding of the risk hidden in the data, hence 
for asset allocation and risk management. These parameters are market characteristics  and provide a simple but 
useful description of the risk, for example in terms of high quantiles such as
Value-at-Risk. Alternatively, these parameters can be
used for model building of the equities in the market such as the GARCH model; see Section~\ref{sec:2}.
%With these motives in mind, we present a survey of the
%tail-indices and scale parameters of 3 sectors, namely ``Energy'',
%``Consumer Staples'' and ``Information Technology'', of the S\&P 500
%index.

\subsection{Hill estimates of lower tail-indices}\label{sec:Hill}
Various estimators of the tail-index $\alpha$  in the model
\eqref{eq:1a} have been proposed in the literature;
see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, de
Haan and Ferreira \cite{haan:ferreira:2006}, Resnick
\cite{resnick:2007}. The most popular among them was introduced  by
Hill \cite{hill1975simple}.
Given a sample $-X_1,\ldots,-X_n$ whose marginal \ds\ satisfies  \eqref{eq:1a}, calculate
the order statistics $X_{(1)}\le \cdots\le X_{(n)}$  and construct the
{\em Hill estimator}:
\beao%\label{eq:Hill_index}
  \hat \alpha_k = \Big(
    {1 \over k} \sum_{i=1}^k \log {X_{(n-i+1)} \over X_{(n-k)}}
    \Big)^{-1}\,.
\eeao
Here $k$  is the number of upper order statistics in the sample used
for the estimation. The estimator $\hat \alpha_k$ is 
a maximum-likelihood estimator of $\alpha$ based on the $k$ upper
order statistics in the pure Pareto model (recall that 
we multiplied the data by minus one)
\beam\label{eq:3}
\P(-X_t>x)= \dfrac{K^\alpha}{x^\alpha}\,,\qquad x>K\,,
\eeam
under the hypothesis that we do not know the (high) threshold value
$K$. The estimator has ``good'' theoretical properties such as 
\asy\ consistency and \asy\ normality. These properties hold under
strict stationarity assumptions on the data; Drees and Rootz\'en 
\cite{drees:rootzen:2010} give perhaps most general conditions for
dependent \seq s and de Haan and Ferreira \cite{haan:ferreira:2006} provide
a complete \asy\ theory in the iid case.
\par
A major problem for Hill estimation is the choice of the number $k$ of
upper order statistics. As  a matter of fact,
if $k$ is too large the order statistics are too close to the center
of the \ds\ of the $-X_t$, leading to a bias
of the estimator. On the other hand, by construction, $\hat \alpha_k$ is an average of $k$ log-differences of the 
data. Therefore, the variance of the estimator is the larger the
smaller $k$. For these reasons, \asy\ theory requires 
to choose $k=k_n$ \st\ $k_n\to\infty$ and $k_n/n\to 0$ as $\nto$. This
fact does not make the estimation of $\alpha$ an  
easy matter: one has to choose a ``small'' value $k$ which is not
``too large''. For practical purposes,
a so-called {\em Hill plot} is recommended where $\hat \alpha_k$ is
plotted for a variety of $k$-values, corresponding to
some high  quantile $X_{(n-k)}$ of the data. Then $k$ is chosen from a
region in the plot where $\hat \alpha_k$ is relatively
stable. For example, in Figure~\ref{fig:1} we have chosen $k=50$ from
a sample of size $n=1304$, corresponding to the 96\%-quantile of the
data. In general, the estimation of the tail-index is an art and
requires some expertise; for some guidance
see Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997},
Resnick \cite{resnick:2007} and Drees et al. 
\cite{drees:resnick:2000}. 
\par
In Figure~\ref{fig:1},
we exhibit 95\% \asy\ confidence bands derived from  the \clt
\beam\label{eq:2}
\sqrt k\, \big(\hat \alpha_k - \alpha\big) \std N(0, \alpha^2)\,,
\eeam
i.e., $\hat \alpha_k$ is \asy ally unbiased and has variance
$\alpha^2/k$. Since $k/n\to 0$ this means that the 
confidence bands are significantly larger than the classical
$1/\sqrt{n}$-rates. This fact is one explanation for the fact that it is
difficult to say something meaningful about the true value of
$\alpha$. There exist various other reasons why one should
not have 100\% trust in the confidence bands shown in
Figure~\ref{fig:1}. Indeed, \eqref{eq:2}
holds under rather subtle {\em second order conditions} on the tail
$\P(X_t>x)$ which cannot be verified on data. However, 
given a theoretical model such as the GARCH, these conditions can be
verified based on the theoretical properties of the model. If they are not satisfied the 
Hill estimator may exhibit significant bias; see Embrechts et
al. \cite{embrechts:klueppelberg:mikosch:1997} and Resnick \cite{resnick:1987} for illustrations 
of this fact leading to so-called ``Hill horror plots''. Moreover, the
Hill estimator is rather sensitive to non-stationarity 
of the data and to dependence. For example, results in Drees \cite{drees:2008}, and Drees and
Rootz\'en \cite{drees:rootzen:2010} show that the \asy\ variance of the
Hill estimator can be significantly larger than in the iid case. 
Since return data are dependent, the \asy\ confidence bands should be
even wider than exhibited in Figure~\ref{fig:1}. 
Again, only under he assumption of a concrete model like GARCH these
confidence bands can be evaluated and therefore the bands shown in Figure~\ref{fig:1}
just show some benchmark which holds in the iid case and under additional conditions on the tail \asy s.
\begin{figure}[htb!]
  \begin{minipage}{1.0\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
    {Energy_lower.pdf}
  \end{minipage}
  \begin{minipage}{1.0\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
    {Consumer_Staples_lower.pdf}
  \end{minipage}
  \begin{minipage}{1.0\linewidth}
    \includegraphics[width=\textwidth, trim={0, 0.8cm, 0, 2cm}, clip]
    {Information_Technology_lower.pdf}
  \end{minipage}
  \caption{\small Hill estimates $\hat \alpha_{50}$ of the lower tail-indices $\alpha$ of
    daily return series in sectors of the S\&P 500
    index. The data span
    from 1 January 2010 to 31 December 2014 and comprise $n=1304$ observations.
The graphs from top to bottom correspond to the ``Energy'',
    ``Consumer Staples'' and ``Information Technology'' sectors.
    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the gray
    triangles above and below it mark the 97.5\% and 2.5\% quantiles
    of its approximate normal distribution; see \eqref{eq:2} and the discussion following it for an 
interpretation.
    The lower and upper red lines mark the medians of the 2.5\% 
    and 97.5\% quantiles, respectively, evaluated from all stocks in the sector.
    The data are taken from {\it Yahoo Finance}; the labels on
    the horizontal axes are Yahoo symbols of the stocks. 
  }\label{fig:1}
\end{figure}

In Figure~\ref{fig:1} we see significant overlap of the confidence intervals of the Hill
estimates of the losses in the ``Energy'' and ``Consumer Staples''
sectors of the S\&P 500 index, as well as those of a 
large portion of losses in the ``Information Technology'' sector.
This fact indicates that the returns in each sector may 
have comparable tail-indices.
\par
Hoga's \cite{hoga:2016} test about the change of extreme quantiles
in a sample
may provide some further insight about how similar these tail-indices are.
Different tail-indices are likely to result in different
extreme quantiles. Nevertheless, changes in the extreme quantiles may also
result from changing scale parameters in the tail. Therefore  we first investigate the scale
parameters of daily stock returns in the same sectors of S\&P 500 before we apply the test.

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
\beao
  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
    0, K^{2\alpha}  \big)\,,
\end{eqnarray*}
where the tail-index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of $\hat K_k$
and $\hat K_k^{\hat\alpha}$ can be constructed.   
Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {CS_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {IT_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $\hat K_k$ (top) and $\hat K_k^{\hat
      \alpha}$ (bottom) on $\log_{10}$-scale of stocks in sectors
    of the S\&P 500 index. The estimates are ordered according to the 
    corresponding estimated $\alpha$-values.
    The points are the estimated values, the bars 
    the \asy\ 95\%-confidence intervals; the confidence bands of the
    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
    shown in Figure~\ref{fig:1}. 
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd row of
Figure~\ref{fig:sectors_parameters}.
\par
Secondly, it appears that there is positive dependence between the
values of $\alpha$ and $K$.
As argued in Section~\ref{sec:pareto_tail}, this is consistent with
the assumption that the return series have Pareto tails on both sides
with the tail parameters on each side independent of those on the
other.

\par
Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
average, the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are  larger
than those in the ``Consumer Staples'' sector.  For a given loss \pro y, 
a larger value of $K$ implies that large losses are more
probable. Thus one can conclude that these two sectors are
considerably riskier than the ``Consumer Staples'' sector. {\red Casper:
we leave this sentence to your discretion.} This is of course a
confirmation of one's economic instinct.

Yet another indication from Figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the dependence  between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, when moving along
a curve of equal preference in the direction of increasing $\alpha$, the parameter $K$ also
increases. So the strong positive dependence seen in the ``Energy''
sector suggests that these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and do not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

\subsection{A test for equal tail-indices based on Hill estimation}
Suppose we have two independent strictly stationary positive series $X_1, \dots, X_n $ and
$Y_1, \dots, Y_n$ with corresponding \df s $F_X$ and $F_Y$ that have power-law tails
with indices $\alpha_X$ and $\alpha_Y$, respectively. 
From \eqref{eq:2} one can deduce
\begin{equation}
  \label{eq:x3}
  \sqrt k

>>>>

\documentclass[11pt,a4]{amsart}
\input{../../ThomasDefs}
\begin{document}
\title{Do return series have power-law tails with the same index?}
\author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}

@@

\input{PaperBody}
